### 2020-02-07 ncov_lancet_ID.tex
editing lancet ID version

We decompose \Ro into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
We use a Bayesian multilevel model tomeasure uncertainties associated with these quantities.
We assess early estimates of \Ro for the 2019-nCoV outbreak across these ranges of uncertainties.

@@

@@

@@

We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23--26, 2020 (Table 1).
As most studies do not report their estimates of the exponential growth rate, we first recalculate the exponential growth rate that correspond to their model assumptions.
We do so by modeling reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
we used gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
we model this estimate as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
We then constructed a family of parameter sets --- which include $r$, $\bar G$, and $\kappa$ --- for each study and used these in a Bayesian multilevel model to build a distribution of pooled estimates for each parameter ($\mu_r$, $\mu_G$, and $\mu_\kappa$).
The pooled estimates, which are represented as probability distributions rather than point estimates, allow us to average across different modeling approaches, while accounting for the uncertainties in the assumptions they make (see Methods).

@@

We reviewed 7 modeling studies of the novel coronavirus outbreak that were published online between January 23 and January 26, 2020 (Table 1).
Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
They vary widely in their statistical methods and in the amount of data used to infer \Ro.
Bedford \textit{et al.} \cite{bedfordncov}, Imai \textit{et al.} \cite{imaincov}, and Riou and Althaus \cite{riouncov} simulated branching process models and compared the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0}).
Read \textit{et al.} \cite{readncov} fitted a deterministic, metapopulation Susceptible-Exposed-Infected-Recovered (SEIR) model to incidence data between January 1 and January 21 from major cities in China and other countries.
Zhao \textit{et al.} \cite{zhaoncov} fitted an exponential growth model to incidence data up to January 22 and inferred \Ro via the Euler-Lotka equation (\eref{euler}).
Liu \textit{et al.} \cite{liuncov} modeled the observed incidence up to January 23 as a function of the observed incidence at previous time steps, the generation-interval distribution, and the basic reproductive number using a Poisson likelihood \citep{forsberg2008likelihood}.
Finally, Majumder and Mandl \textit{et al.} \cite{majumderncov} fitted the Incidence Decay and Exponential Adjustment (IDEA) model \citep{fisman2013idea} to incidence data up to January 26, which is equivalent to fitting an exponential growth model and assuming a fixed generation-interval distribution.
Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.

For each study $i$, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Rx{0i} and the assumed values of $\bar G_i$ and $\kappa_i$ and calculating the exponential growth rate $r_i$ via the inverse of \eref{gamma}:
\begin{linenomath*}
\begin{equation}
r_i = \frac{\Rx{0i}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
\end{equation}
\end{linenomath*}
This allows us to approximate the probability distributions of the estimated exponential growth rates by each study;
uncertainties in the probability distributions that we calculate for the estimated exponential growth rates will reflect the methods and assumptions that the studies rely on.

We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution:
\begin{linenomath*}
\begin{equation}
\begin{aligned}
r_i &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\mu_r^2/\sigma_r^2),\\
\bar{G}_i &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\mu_G^2/\sigma_G^2),\\
\kappa_i &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\mu_\kappa^2/\sigma_\kappa^2),\\
\end{aligned}
\end{equation}
\end{linenomath*}
where $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations in $r$, $\bar G$, and $\kappa$, respectively.
We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step.
Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.
We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
As we add more models, we expect the pooled estimates to become sharper even when the models no longer add more information.
Thus, the pooled estimator should be interpreted with care.

>>>>

We decompose \Ro into three key quantities: the exponential growth rate $r$, the mean generation interval $\bar G$, and the generation-interval dispersion $\kappa$.
We use a Bayesian multilevel model to construct pooled estimates and measure uncertainties associated with these quantities.
We assess early estimates of \Ro for the 2019-nCoV outbreak using the pooled estimates.

@@

\subsection*{Evidence before this study}

7 analyses of the 2019-nCoV outbreak were published online as pre-prints between January 23--26, 2020.
Their modeling approaches varied widely and estimates of uncertainties in \Ro ranged from 1.5 to 7.1.

This section should include a description of all the evidence that the authors considered before undertaking this study. Authors should briefly state: the sources (databases, journal or book reference lists, etc) searched; the criteria used to include or exclude studies (including the exact start and end dates of the search), which should not be limited to English language publications; the search terms used; the quality (risk of bias) of that evidence; and the pooled estimate derived from meta-analysis of the evidence, if appropriate

\subsection*{Added value of this study}

Authors should describe here how their findings add value to the existing evidence

\subsection*{Implications of all the available evidence}

Authors should state the implications for practice or policy and future research of their study combined with existing evidence.Research in context panels should not contain references; key studies mentioned here should be referenced in the main text.

@@

We gathered information on estimates of \Ro and their assumptions about the underlying generation-interval distributions from 7 articles that were published online between January 23--26, 2020 (Table 1).
Five studies \citep{liuncov, majumderncov, readncov, riouncov, zhaoncov} were uploaded to pre-print servers (bioRxiv, medRxiv, and SSRN); one report was posted on the web site of Imperial College London \citep{imaincov}; and one report was posted on nextstrain.org \citep{bedfordncov}.
They vary widely in their modeling approaches: a branching process model \citep{bedfordncov, imaincov, riouncov}, a deterministic Susceptible-Exposed-Infected-Recovered (SEIR) model \citep{readncov}, an exponential growth model \citep{zhaoncov}, a Poisson offspring distribution model \citep{liuncov}, and the Incidence Decay and Exponential Adjustment (IDEA) model \citep{majumderncov}.
Four studies estimated \Ro by directly fitting their models to incidence data \citep{readncov, zhaoncov, liuncov, majumderncov}.
The remaining three studies estimated \Ro by comparing the predicted number of cases from their models with the estimated number of total cases by January 18 (between 1,000 and 9,7000 \citep{imaincov0})
Some of these studies have now been published in peer-reviewed journals \citep{riou2020pattern, zhao2020preliminary} or have been updated with better uncertainty quantification \citep{readncov2}.

As most studies do not report their estimates of the exponential growth rate, we first recalculate the exponential growth rate that correspond to their model assumptions.
We do so by modeling reported distributions of the reproductive number \Ro, the mean generation interval $\bar G$, and the generation-interval dispersion parameter $\kappa$ with appropriate probability distributions;
we used gamma distributions to model values reported with confidence intervals and uniform distributions to model values reported with ranges.
For example, Study 3 estimated $\Ro = 2.92$ (95\% CI: 2.28--3.67);
we model this estimate as a gamma distribution with a mean of 2.92 and a shape parameter of 67, which has a 95\% probability of containing a value between 2.28 and 3.67 (see Table 2 for a complete description).
For each study $i$, we construct a family of parameter sets by drawing 100,000 random samples from the probability distributions (Table 2) that represent the estimates of \Rx{0i} and the assumed values of $\bar G_i$ and $\kappa_i$ and calculate the exponential growth rate $r_i$ via the inverse of \eref{gamma}:
\begin{linenomath*}
\begin{equation}
r_i = \frac{\Rx{0i}^{\kappa_i} - 1}{\kappa_i \bar{G}_i}.
\end{equation}
\end{linenomath*}
This allows us to approximate the probability distributions of the estimated exponential growth rates by each study;
uncertainties in the probability distributions that we calculate for the estimated exponential growth rates will reflect the methods and assumptions that the studies rely on.

We construct pooled estimates for each parameter ($r$, $\bar G$, and $\kappa$) using a Bayesian multilevel modeling approach, which assumes that the parameters across different studies come from the same gamma distribution:
\begin{linenomath*}
\begin{equation}
\begin{aligned}
r_i &\sim \gammdist(\mathrm{mean}=\mu_r, \mathrm{shape}=\mu_r^2/\sigma_r^2),\\
\bar{G}_i &\sim \gammdist(\mathrm{mean}=\mu_G, \mathrm{shape}=\mu_G^2/\sigma_G^2),\\
\kappa_i &\sim \gammdist(\mathrm{mean}=\mu_\kappa, \mathrm{shape}=\mu_\kappa^2/\sigma_\kappa^2),\\
\end{aligned}
\end{equation}
\end{linenomath*}
where $\sigma_r$, $\sigma_G$, and $\sigma_\kappa$ represent between-study standard deviations in $r$, $\bar G$, and $\kappa$, respectively.
We account for uncertainties associated with $r_i$, $\bar G_i$ and $\kappa_i$ (and their correlations), by drawing a random set from the family of parameter sets for each study at each Metropolis-Hastings step.
Since the gamma distribution does not allow zeros, we use $\kappa=0.02$ instead for Study 7.
We note that this approach does not account for non-independence between the parameter estimates made by different modelers.
As we add more models, the pooled estimates can become sharper even when the models no longer add more information.
Thus, the pooled estimator should be interpreted with care.

@@

@@

\section*{Contributors}

SWP and JD developed the statistical framework. 
SWP reviewed the published literature.
SWP performed the analysis. 
SWP, BMB, and JD created the figures. 
SWP and JD wrote the first draft.
All authors contributed to the writing and approval of the final report.

\section*{Declaration of interests}

We declare no competing interests.
